Scalar Feedback based Joint Time-Frequency Precoder interpolation for MIMO-OFDM Systems

### 2019-10-14 paper/main.tex
Add changes

\subsection{Differential Quantization - Channel Tracking}
\label{quantiz}
For a channel that varies slowly with time, ADPM is an effective
method for tracking parameters over time using very few bits (often
just one bit). For example, while tracking a process $x_n$, at time
$n$, the one-bit jump $\beta_{n} = Q(x_{n} - \hat{x}_{n-1})$, where
$x_n$ is the unquantized value and $\hat{x}_{n-1}$ is the unquantized
(accurate) value, the quantization function $Q(x)$ is given as
\begin{equation}
  Q(x)=\begin{cases}
    1, & \text{if $x>0$}.\\
    -1, & \text{otherwise}.
  \end{cases}
\end{equation}
Thus, the quantied vector $\hat{x}_n$ evolves with time as
\begin{equation}
\hat{x}_{n} = \hat{x}_{n-1} + \beta_{n}\Delta_{n} \mbox{ where }
\label{delta_eqn}
\Delta_{n} = \begin{cases}
    M \Delta_{n-1}, & \text{if $\beta_{n} = \beta_{n-1}$}\\
    \Delta_{n-1}/m , & \text{if $\beta_{n} \neq \beta_{n-1}$}.
  \end{cases}
\end{equation}
Here we initialize delta using $\Delta_1 = |x_{2}-\hat{x}_1|$.

@@

\section{Simulation and discussion}
\label{section3}
We present some simulations of the scalar feedback based joint
time-frequency predictive quantization outlined in the previous section. performed simulations to analyse the performance of the proposed quantization and interpolation method for slowly time-varying MIMO OFDM channel. Channel is simulated using the typical Jakes Model for both ITU Vehicular-A channel and ITU Pedestrian Channel. Performance is measured by averaging over 10 channel realizations with 100 simulations each. The normalised doppler frequency $f_{norm} = f_{max}.T_s$ is set to $3.5\times10^{-4}$ with symbol time period $T_d = 5\times10^{-8} s$. For both vehicular and pedestrian channel, $N_T=4$ and $N_R=2$. Therefore, the total number of parameters are $N_{R}(2N_{T} - N_{R}) = 12$ with 5$\theta$’s and 7$\phi$’s. There are 1024 subcarriers in Vehicular channel and 64 subcarriers in the Pedestrian Channel. We take s = 33/9 for Vehicular and pedestrian channel respectively i.e. 32/8 equally spaced subcarriers with indices 33t/9t, t = 0,1,$, .. , 31 and 0,1, .. ,7 respectively.

\subsection{Initial Quantization}
For the first and the second simulations/time instances we use of $\textbf{B}$ = 10 bits for quantizing each parameter for effective quantization/initialization. Therefore, the number of bits for each subcarrier = $10\times12$ = 120. Quantization is performed as given in Section~\ref{section2}.

Simulation: BER is calculated by averaging over all the simulations. For Vehicular Channel, 1024 subcarriers are used whereas 64 subcarriers are used in the Pedestrian Channel.

Initialization is performed by using $10\times12$ bits for each subcarrier for the initial two time frames. Later, the quantization takes place in the time-domain using one bit for each parameter. Since the bit budget is low and we are considering slowly varying channels, therefore, either theta or phi parameters are sent alternatingly in the time domain while keeping the unsent parameters equal to their previous values as show in the Figure \ref{fig:adpm-fig}. Here $\Theta$ is the the collection of all $\theta$'s and $\Phi$ is the collection of all $phi$'s. This brings down the bit budget to average of 5(Size of $\Theta$) and 7(Size of $\Phi$) i.e. 6 bits average per fed back subcarrier.

{ADPM Design}: The decrement/decay rate is given by $\bm$ and increment/growth rate is given by $\bM$ in the Equation~\ref{delta_eqn} . The ratio of $m/M$ is fixed at 2.5 i.e. higher decay than increment of the delta values. This way, quantized values are able to follow the curve more precisely. Also, need to describe the initialization of the Delta values required in the quantization process. Need to cite that paper.

>>>>

\subsection{Differential quantization and channel tracking}
\label{quantiz}
For channels that vary slowly with time, adaptive differential
quantization is an effective method for tracking parameters over time
using very few bits (often just one bit). For example, while tracking
a slowly varying discrete random process $x_n$ at time $n$, the one
bit can be used to determine the direction in which to move in order
to track the parameter. This can be denoted by
$\beta_{n} = \mbox{sign}(x_{n} - \hat{x}_{n-1})$, where $x_n$ is the
unquantized value and $\hat{x}_{n-1}$ is the unquantized (accurate)
value. Further, the step size for the move can be adapted based on the
speed of channel variation as well the speed for convergence. If
$\Delta_n$ is the step size, we define our adaptive quantizer as
\begin{equation}
\hat{x}_{n} = \hat{x}_{n-1} + \beta_{n}\Delta_{n} \mbox{ where }
\label{delta_eqn}
\Delta_{n} = \begin{cases}
    m \Delta_{n-1}, & \text{if $\beta_{n} = \beta_{n-1}$}\\
    \Delta_{n-1}/M , & \text{if $\beta_{n} \neq \beta_{n-1}$}.
  \end{cases}
\end{equation}
where $M$ and $m$ are the jumps used during the positive and negative
moves. We note that, in general, $m = M$, however, we have found that
using $M > m$ seems to yield better performance. We initialize
$\Delta_1$ as $\Delta_1 = |x_{2}-\hat{x}_1|$. We note that this is
performed separately for each of the quantized parameters.

@@

\section{Simulation and discussion}
\label{section3}
We present some simulations of the scalar feedback based joint
time-frequency predictive quantization outlined in the previous
section. In particular, we analyze the performance of the proposed
quantization and interpolation method for time-varying MIMO-OFDM
systems. The channel is simulated using the typical Jakes Model for
both the ITU Vehicular-A channel and ITU Pedestrian channel power
delay profiles CITE. The BER achieved by the proposed predictive
quantization based precoder is compared against the ideal precoder for
100 channel evolutions in each run and averaged. The normalised
Doppler is assumed to be $f_dT_s = 3.5\times10^{-4}$, with symbol time
period $T_s = 5\times10^{-8} s$. For both the Vehicular and Pedestrian
channel profiles, we consider $N_T=4$ and $N_R=2$. Therefore, the
total number of parameters that determine the precoder are
$N_{R}(2N_{T} - N_{R}) = 12$, with 5$\theta_l$'s and
7$\phi_i$'s. For the Vehicular channel profile, we assume that the
number of subcarriers is 1024, while for the Pedestrian channel
profile it is 64. We take the spacing between fed back subcarrier
indices $p$ to be $33$ and $9$ for the Vehicular
and Pedestrian channel profiles respectively (i.e. for the Vehicular
profile, the subcarriers indices fed back would be $0, 33, 66, \ldots
1023$ and for the Pedestrian, $0, 9, 18, \ldots 63$).

For the first and the second simulation time instances, since the
precoder is not available at the transmitter, we use 10 bits for
quantizing each parameter for effective representation of the
precoder. Therefore, the number of bits for each subcarrier =
$10\times12$ = 120. Quantization is performed as given in
Section~\ref{section2}. We note that this is not significant, since,
amortized over long durations, this number becomes small. Subsequent
quantization takes place in the time-domain using one bit for each
parameter. Since the bit budget is small, if the channel varies
sufficiently slowly, either the $theta_l$ or the $phi_i$ parameters
are fed back alternately, while retaining the unsent parameters as
being to their previous values, as shows in
Fig.~\ref{fig:adpm-fig}. Here, $\Theta$ is the the collection of all
$\theta_l$s and $\Phi$ is the collection of all $\phi_i$s. This brings
down the bit budget to an average of 5 bits(the length of $\Theta$)
and 7 bits (the length of $\Phi$) over alternate feedback instances,
thereby resulting in an average feedback budget of 5 bits.

To enable adaptive tracking of the channel parameters, we start with
an initialization of the step size for each parameter, as discussed in
Equation~\ref{delta_eqn}, with $m = 0.56, M = 1.4$ (the choice is
motivated by the results in~\ref{5671092}), and track parameters with
the temporal evolution of the channel. We have found that using
$M / m = 2.5$ yields optimum performance in terms of channel tracking.
